---
title: "Human Development Index (HDI) Analysis "
subtitle: "- What a good development entails"
author:
- Jack Gleeson 480382724
- Jelly Dongyi Guo 470033070
- Mayira Noori 500473922
- Yasheita Roy 480358833
date: "University of Sydney | MATH1115 Group 22 | Sept. 2021"
output:
  html_document:
    code_download: true # embed .Rmd file
    toc: true  # table of contents
    toc_depth: 3
    toc_float: true # toc is always visible when scrolled
    number_sections: false 
    df_print: paged # format data frame outputs automatically
    theme: sandstone # theme the doc
    highlight: tango # theme syntax colouring
---

```{css newstyles, echo=FALSE}
body {
  color: black;
}
```

# Executive Summary
- The aim of this report is to ...
- The main discoveries are ...

# Full Report

## Initial Data Analysis (IDA)

Summary (suggesstion):

- The data came from ...
- The data is/is not valid because ...
- Possible issues include ...
- Each row represents ...
- Each column represents ...

## Package Preparation

```{r packages, message=FALSE, warning=FALSE, results='hide'}
# Required Packages
library(tidyverse)
library(ggpubr)
library(readxl)
library(plotly)
library(gapminder)
library(readr)
```

## Research Question 1

> **How has global HDI changed by geographical region in recent years (2010-2019) based on gender?**

```{r} 
data <- read.csv("data/d1.csv")
ggplot(data, aes(fill=Geographic.Region, y=HDI..Male., x= as.factor(Year..1995.2019.)))+ 
geom_bar(position="dodge", stat="identity")+ 
xlab("Year 2010-2019") + 
ylab("HDI male") 

data <- read.csv("data/d1.csv") 
ggplot(data, aes(fill=Geographic.Region, y=HDI..female., x= as.factor(Year..1995.2019.)))+ 
geom_bar(position="dodge", stat="identity")+ 
xlab("Year 2010-2019") + 
ylab("HDI female") 
```

The two bar plots derived from this particular research question utilise four variables of Geographic region (as defined by the United Nations Development Programme), Years 2010-2019 and Male/Female Human Development Index(HDI). The goal of this research question is to identify and explain gender based trends as correlated to geographical region in the most recent decade 2010-2019. World data was also used as an average baseline to show the scope of disparity between high-performing and low performing regions. 

Initially, from these 2 plots we can see increases in both male and female HDI in all geographic regions across the last decade, suggesting improvements to overall quality of life. Europe and central Asia shows the most significant year to year increase in both Male and female, while Sub-Saharan Africa shows the least amount of growth. In relation to differences based on gender, it is clearly evident that male HDI is significantly greater than female HDI, even in the higher performing regions of 1) Europe and Central Asia and 2) Latin America and the Caribbean indicating a significant gap in life quality between men and women. It must be noted that although it is a strong indicator, a reliance on HDI for identifying gender disparity can be misleading (Al-Hilani, H., 2012.) 

## Data Prepartion for Research Question 2 and 3

The data defined here will be used in research question 2 and 3:

```{r message=FALSE, warning=FALSE, results='hide'}

dat_path <- "data/data.xlsx"

# Table 1 - HDI and its Components
hdi_and_components <- read_excel(dat_path, sheet = "Table 1")[-c(75,129,167), c(1,2,3,5,7,9,11)][8:196,]
colnames(hdi_and_components) <- c("Rank", "Country", "HDI", "LifeExp", "ExpSchlYrs","MeaSchlYrs","GNIpcap")
hdi_and_components %>% 
  mutate(Rank = as.numeric(Rank)) %>% 
  mutate(HDI = as.numeric(HDI)) %>% 
  mutate(LifeExp = as.numeric(LifeExp)) %>% 
  mutate(ExpSchlYrs = as.numeric(ExpSchlYrs)) %>% 
  mutate(MeaSchlYrs = as.numeric(MeaSchlYrs)) %>% 
  mutate(GNIpcap = as.numeric(GNIpcap)) %>% 
  mutate(GNIpcap = GNIpcap / 1000) %>% 
  mutate_if(is.numeric, round, digits=2) -> hdi_and_components

# Table 11 - Work and Employments
employment_status <- read_excel(dat_path, sheet = "Table 11")[-c(75,129,167), c(1,2,3,5,7,9,11,13,15)][9:197,]
colnames(employment_status) <- c("Rank", "Country", "EmpRatio", "LbrFPr", "EmpAgri","EmpServ","UEmpTotal","UEmpYouth","UEmpYNSchl")
employment_status %>% 
  mutate(Rank = as.numeric(Rank)) %>% 
  mutate(EmpRatio = as.numeric(EmpRatio)) %>% 
  mutate(EmpAgri = as.numeric(EmpAgri)) %>% 
  mutate(EmpServ = as.numeric(EmpServ)) %>% 
  mutate(LbrFPr = as.numeric(LbrFPr)) %>% 
  mutate(UEmpTotal = as.numeric(UEmpTotal)) %>% 
  mutate(UEmpYouth = as.numeric(UEmpYouth)) %>% 
  mutate(UEmpYNSchl = as.numeric(UEmpYNSchl)) %>% 
  mutate_if(is.numeric, round, digits=2) -> employment_status
```

## Research Question 2

> **Among all the HDI components, which category has the most strong reflection with the HDI?**


To find the category with the most strong reflection with the HDI index, we are looking for a linear relationship between them with the distribution of data as tight as possible. In this case, all of the categories have a liner model there with HDI:

```{r}
hdi_and_components %>% 
  select(HDI,LifeExp) %>% 
  drop_na %>%
  ggplot(aes(HDI, LifeExp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Index") +
  ylab("Life Expectancy at Birth (Years)") +
  ggtitle("HDI and Life Expectancy at Birth") +
  theme_minimal() +
  geom_blank() -> plot_LifeExp

hdi_and_components %>% 
  select(HDI,ExpSchlYrs) %>% 
  drop_na %>%
  ggplot(aes(HDI, ExpSchlYrs)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Index") +
  ylab("Expected School Years (Years)") +
  ggtitle("HDI and Expected School Years") +
  theme_minimal() +
  geom_blank() -> plot_ExpSchlYrs

hdi_and_components %>% 
  select(HDI,GNIpcap) %>% 
  drop_na %>%
  ggplot(aes(HDI, GNIpcap)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Index") +
  ylab("Gross national income per capita ($1k)") +
  ggtitle("HDI and GNI per capita") +
  theme_minimal() +
  geom_blank() -> plot_GNI

hdi_and_components %>% 
  select(HDI,MeaSchlYrs) %>% 
  drop_na %>%
  ggplot(aes(HDI, MeaSchlYrs)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Index") +
  ylab("Mean School Years (Years)") +
  ggtitle("HDI and School Years") +
  theme_minimal() +
  geom_blank() -> plot_MeaSchlYrs

hdi_and_components %>% 
  select(Rank,LifeExp) %>% 
  drop_na %>%
  ggplot(aes(Rank, LifeExp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Rank (Lower is Better)") +
  ylab("Life Expectancy at Birth (Years)") +
  ggtitle("HDI and Life Expectancy at Birth") +
  theme_minimal() +
  geom_blank() -> plot_LifeExp_rank

hdi_and_components %>% 
  select(Rank,ExpSchlYrs) %>% 
  drop_na %>%
  ggplot(aes(Rank, ExpSchlYrs)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Rank (Lower is Better)") +
  ylab("Expected School Years (Years)") +
  ggtitle("HDI and Expected School Years") +
  theme_minimal() +
  geom_blank() -> plot_ExpSchlYrs_rank

hdi_and_components %>% 
  select(Rank,GNIpcap) %>% 
  drop_na %>%
  ggplot(aes(Rank, GNIpcap)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Rank (Lower is Better)") +
  ylab("Gross national income per capita ($1k)") +
  ggtitle("HDI and GNI per capita") +
  theme_minimal() +
  geom_blank() -> plot_GNI_rank

hdi_and_components %>% 
  select(Rank,MeaSchlYrs) %>% 
  drop_na %>%
  ggplot(aes(Rank, MeaSchlYrs)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Rank (Lower is Better)") +
  ylab("Mean School Years (Years)") +
  ggtitle("HDI and School Years") +
  theme_minimal() +
  geom_blank() -> plot_MeaSchlYrs_rank
  
plot_index <- ggarrange(plot_LifeExp, plot_ExpSchlYrs, plot_MeaSchlYrs, plot_GNI, nrow = 2, ncol = 2)
plot_rank <- ggarrange(plot_LifeExp_rank, plot_ExpSchlYrs_rank, plot_MeaSchlYrs_rank, plot_GNI_rank, nrow = 2, ncol = 2)
```

Now we can peak the distribution and the linear model easily with the scatter plot of both the HDI index and the rank:

```{r}
plot_index
plot_rank
```

By observing the linear model, especially correspondingly with both the rank and index in a different angle, Expected School Years has both tighter data distribution and less distance between maximum and minimum value, especially in range from 0.6 to 0.8 in HDI value and from 50 to 125 in HDI rank. GNI does have a more tight model but it has so many extremely high values and it makes it hard to describe or predict the HDI performance simply because of it.This can be proven more by looking into the standard derivation:

```{r}
hdi_and_components %>% 
  select(LifeExp, ExpSchlYrs, MeaSchlYrs, GNIpcap) %>% 
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  summarise(sd = round(sd(value,na.rm = TRUE),digits = 2))
```

Thus, Expected School Years has the most strong relationship with the HDI.

## Research Question 3

> **How does the employment status of each country related to their HDI rank?**

In order to analyse how the employment situation of each country can relate to their HDI rank, we can use the `plot()` function to glimpse an overview:

```{r}
employment_status %>% 
  keep(is.numeric) %>% 
  sample_frac(0.1) %>% 
  plot()
```

It seems like there are interesting behaviors happening on `EmpAgri` and `EmpServ`, Let's zoom in:

```{r}
employment_status %>% 
  select(Rank,EmpAgri) %>% 
  drop_na %>%
  ggplot(aes(Rank, EmpAgri)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Rank (Lower is Better)") +
  ylab("Employment in Agriculture (%)") +
  ggtitle("HDI Rank and Employment in Agriculture") +
  theme_minimal() +
  geom_blank() -> plot_EmpAgri

employment_status %>% 
  select(Rank,EmpServ) %>% 
  drop_na %>%
  ggplot(aes(Rank, EmpServ)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Rank (Lower is Better)") +
  ylab("Employment in Services (%)") +
  ggtitle("HDI Rank and Employment in Services") +
  theme_minimal() +
  geom_blank() -> plot_EmpServ

ggarrange(plot_EmpAgri, plot_EmpServ, nrow = 2, ncol = 1)
```

It is quite clear and direct: The country ranked higher in the HDI will engage a obviously high rate of employment in services and low rate of employment in agriculture, stepping into the third industry is an essential key to a better development.

But how about the overall employment and unemployment, how would they affect the HDI?

```{r}
employment_status %>% 
  select(Rank,EmpRatio) %>% 
  drop_na %>%
  ggplot(aes(Rank, EmpRatio)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Rank (Lower is Better)") +
  ylab("Employment (%)") +
  ggtitle("HDI Rank and Employment") +
  theme_minimal() +
  geom_blank() -> plot_EmpRation

employment_status %>% 
  select(Rank,UEmpTotal) %>% 
  drop_na %>%
  ggplot(aes(Rank, UEmpTotal)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlab ("HDI Rank (Lower is Better)") +
  ylab("Unemployment (%)") +
  ggtitle("HDI Rank and Unemployment") +
  theme_minimal() +
  geom_blank() -> plot_EmpUEmpTotal

ggarrange(plot_EmpRation, plot_EmpUEmpTotal, nrow = 2, ncol = 1)
```

Based on how horizontal the linear line is, the conclusion is despite higher HDI-ranked countries seems to like have a higher employment ration and lower unemployment, there is still no connection that having more people employed or unemployed will affect the HDI ranking. This also applies to relating variables of these 2 rations like Lab our Force Participation Rate and Youth unemployment (or school), which can be easily observed in the `plot()` before.

## Research Question 4

> **Is there a difference in 'Workforce Participation Rate (WPR)' between males and females in the four different HDI catgeories?**

The UN has cartgeorised the HDI value of countries into 4 broad classifications that provide an indication of the countries development as shown below:

- Very High Human Development = HDI Value > 0.898
- High Human Development = HDI Value > 0.753
- Medium Human Development = HDI Value > 0.631
- Low Development = HDI Value > 0.513

Through the use of a violin plot we were able to visualise the WPR (%) of males and females in relation the HDI value of the country, which brought to forefront clear trends. 

There are clear seperations between the WPR of the genders and the HDI class of the country namely one being, female participation rate on average remains lower than male participation rates, despite the HDI category. This may be caused by remenants of societal values that did not encourage or rather presented barriers of entry for females into the workforce. Futhermore, it is often expected for females to also be primary caregivers which may reduce the tendancy for WFP across the board (The gender gap in employment: What's holding women back?, 2021)

Very High HDI countries tend to have a WPR with a smaller distribution than that of Low HDI countries in both genders, and is clearly depicted as the size of the violins elongate as HDI value decreases. Higher HDI countries tend to be more bulbous indicating a certain level of uniformity within the living standards compared to the larger disparity in lower HDI countries.

In summary it is clearly evident that WPR have unique characteristics for each HDI category that are evident across genders. Futhermore the participation rate of genders do not tend to be a good indicator of the countries HDI value or category.

```{r}
LF_data <- read_csv("data/Labour Force Data.csv")
head(LF_data)
```

```{r}
mv <- LF_data %>%
  ggplot( aes(`HDI Value`, `Male Labour Force Particiaption Rate`, fill= `HDI Category`)) +
  geom_violin() +
  theme_bw()

mv
```

```{r}
fv <- LF_data %>%
  ggplot( aes(`HDI Value`, `Female Labour Force Participation Rate`, fill= `HDI Category`)) +
  geom_violin() +
  theme_bw()

fv
```

# References

- Al-Hilani, H., 2012. HDI as a measure of human development: A better index than the income approach. Journal of Business and Management, vol. 2, no. 5, pp.24-28.

- Ilo.org. 2021. The gender gap in employment: What's holding women back?. [online] Available at: <https://www.ilo.org/infostories/en-GB/Stories/Employment/barriers-women#pressure> [Accessed 5 September 2021].

